### 2021-09-12 documentation_and_analysis/draftbook.Rmd
Adding raw data + setting up project

---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(readODS)
library(httr)
```

```{r}
read_ods("Downloads/electric-vehicle-charging-device-statistics-july-2021.ods")
```

```{r}
APE <- read_ods("Downloads/APE_site_data_tables.ods", sheet = 2)
```

```{r}
NO2 <- read_ods("Downloads/NO2_tables.ods")
```

```{r}
registry <- GET("http://chargepoints.dft.gov.uk/api/retrieve/registry/format/csv/") %>% 
  content()
```

```{r}

```

```{r}
type <- GET("http://chargepoints.dft.gov.uk/api/retrieve/type/format/csv/") %>% 
  content()
```

```{r}
read_csv("Documents/GitHub/ev_climate_change_project/raw_data/no2_by_grid_2019.csv")
```

### 2021-09-13 documentation_and_analysis/draftbook.Rmd
EV Geospatial Working

@@
 library(tidyverse)
+library(janitor)
 library(readODS)
 library(httr)
+library(here)
+library(data.table)
+library(sf)
+library(leaflet)
 ```
@@
 ```{r}
-read_ods("Downloads/electric-vehicle-charging-device-statistics-july-2021.ods")
+read_ods(here("raw_data/electric-vehicle-charging-device-statistics-june-2021.ods"))
 ```
@@
 ```{r}
-APE <- read_ods("Downloads/APE_site_data_tables.ods", sheet = 2)
+getwd()
 ```
@@
 ```{r}
-NO2 <- read_ods("Downloads/NO2_tables.ods")
+APE <- read_ods("raw_data/APE_site_data_tables.ods", sheet = 2)
+```
+
+```{r}
+NO2 <- read_ods("raw_data/NO2_tables.ods")
 ```
@@
   content()
-```
-
-```{r}
-
 ```
@@
 ```
+
+```{r}
+air_pollution <- GET("https://uk-air.defra.gov.uk/data/sos/service?service=AQD&version=1.0.0&request=GetObservation&temporalFilter=om:phenomenonTime,After,yyyy-mm-ddT00:00:00.000Z") 
+```
+
+```{r}
+air_pollution <- rbindlist(air_pollution$content, fill = TRUE) 
+```
+
+```{r}
+air_pollution <- unnest(air_pollution)
+```
+
+# How many electric vehicles are on the road across the UK?
+
+```{r}
+# Reading in and skipping first 5 rows
+uk_ev <- read_ods(here("raw_data/ev_by_la.ods"), skip = 5)
+# Making row 1 the variable name
+names(uk_ev) <- uk_ev[1,] 
+# Removing row 1 
+uk_ev <- uk_ev[-1,] %>% 
+  clean_names()
+```
+
+```{r}
+# loading in shape file 
+uk_shape_file <- st_read(here("raw_data/Local_Authority_Districts__April_2019__UK_BFE_v2-shp/Local_Authority_Districts__April_2019__UK_BFE_v2.shp")) %>%
+  clean_names() %>% 
+  st_simplify(dTolerance = 1000) %>%
+  st_transform("+proj=longlat +datum=WGS84") %>% 
+  select(lad19cd, long, lat, geometry)
+```
+
+```{r}
+# Joining uk_ev + shape file 
+uk_ev_map <- uk_ev %>% 
+  left_join(uk_shape_file, by = c("ons_la_code_apr_2019" = "lad19cd")) %>% 
+  drop_na() %>% 
+  mutate(across(c(x2021_q1:x2011_q4), as.numeric))
+```
+
+```{r}
+    pal <- colorBin("Greens", domain = uk_ev_map$x2021_q1, bins = 10)
+
+    uk_ev_map_labels <- sprintf(
+      uk_ev_map$region_local_authority_apr_2019_3, uk_ev_map$x2021_q1)
+```
+
+```{r}
+leaflet(uk_ev_map) %>% 
+  setView(lng = -4.2026, lat = 55.8, zoom = 4.7, options = list()) %>%
+  addProviderTiles(providers$CartoDB.Positron) %>% 
+  addPolygons(data = uk_ev_map$geometry,
+             # fillColor = ~pal(x2021_q1),
+              weight = 0.5,
+              opacity = 0.9, 
+              color = "green",
+              fillOpacity = 0,
+              highlightOptions = highlightOptions(color = "white", weight = 2,
+                                                      bringToFront = TRUE),
+              label = uk_ev_map_labels,
+                  labelOptions = labelOptions(
+                    style = list("font-weight" = "normal", padding = "3px 8px"),
+                    textsize = "15px",
+                    direction = "auto")) %>% 
+  addLegend(pal = pal, values = ~density, opacity = 0.7, title = NULL,
+                position = "bottomright")
+```
+
+```{r}
+uk_ev_map$x2021_q1 %>% 
+  range()
+```

### 2021-09-13 documentation_and_analysis/draftbook.Rmd
rbind of no2 data

@@
 # Reading in and skipping first 5 rows
-uk_ev <- read_ods(here("raw_data/ev_by_la.ods"), skip = 5)
+uk_ev <- read_ods(here("raw_data/ev_by_la.ods"), sheet = 2, skip = 5)
 # Making row 1 the variable name
@@
   drop_na() %>% 
-  mutate(across(c(x2021_q1:x2011_q4), as.numeric))
+  mutate(across(c(x2021_q1:x2011_q4), as.numeric)) %>% 
+  st_as_sf()
 ```
@@
 ```{r}
-    pal <- colorBin("Greens", domain = uk_ev_map$x2021_q1, bins = 10)
+    pal <- colorBin("Greens", domain = uk_ev_map$x2021_q1, bins = c(0, 500, 1000, 2500, 5000, 10000, 15000))
 
     uk_ev_map_labels <- sprintf(
-      uk_ev_map$region_local_authority_apr_2019_3, uk_ev_map$x2021_q1)
+      "<strong>%s</strong><br/>%g Electric Vehicles",
+      uk_ev_map$region_local_authority_apr_2019_3, uk_ev_map$x2021_q1) %>% 
+      lapply(htmltools::HTML)
 ```
@@
 ```{r}
-leaflet(uk_ev_map) %>% 
+uk_ev_map %>% 
+  leaflet() %>% 
   setView(lng = -4.2026, lat = 55.8, zoom = 4.7, options = list()) %>%
   addProviderTiles(providers$CartoDB.Positron) %>% 
-  addPolygons(data = uk_ev_map$geometry,
-             # fillColor = ~pal(x2021_q1),
-              weight = 0.5,
+  addPolygons(fillColor = ~pal(x2021_q1),
+    weight = 0.1,
     opacity = 0.9, 
-              color = "green",
-              fillOpacity = 0,
-              highlightOptions = highlightOptions(color = "white", weight = 2,
+    color = "black",
+    fillOpacity = 0.8,
+    highlightOptions = highlightOptions(color = "green", weight = 2,
                                         bringToFront = TRUE),
@@
       direction = "auto")) %>% 
-  addLegend(pal = pal, values = ~density, opacity = 0.7, title = NULL,
+  addLegend(pal = pal, values = ~x2021_q1, opacity = 0.7, title = NULL,
             position = "bottomright")
 ```
-
-```{r}
-uk_ev_map$x2021_q1 %>% 
-  range()
-```

### 2021-09-14 documentation_and_analysis/draftbook.Rmd
no2 rowbind working

@@
 ```
+
+# Row binding grid NO2 data
+
+```{r}
+no2_2010 <- read_csv(here("raw_data/no2_by_grid_2010.csv"), skip = 6, 
+  col_names = c("uk_grid_code", "x", "y", "no2"))
+no2_2011 <- read_csv(here("raw_data/no2_by_grid_2011.csv"), skip = 6, 
+  col_names = c("uk_grid_code", "x", "y", "no2"))
+no2_2012 <- read_csv(here("raw_data/no2_by_grid_2012.csv"), skip = 6, 
+  col_names = c("uk_grid_code", "x", "y", "no2"))
+no2_2013 <- read_csv(here("raw_data/no2_by_grid_2013.csv"), skip = 6, 
+  col_names = c("uk_grid_code", "x", "y", "no2"))
+no2_2014 <- read_csv(here("raw_data/no2_by_grid_2014.csv"), skip = 6, 
+  col_names = c("uk_grid_code", "x", "y", "no2"))
+no2_2015 <- read_csv(here("raw_data/no2_by_grid_2015.csv"), skip = 6, 
+  col_names = c("uk_grid_code", "x", "y", "no2"))
+no2_2016 <- read_csv(here("raw_data/no2_by_grid_2016.csv"), skip = 6, 
+  col_names = c("uk_grid_code", "x", "y", "no2"))
+no2_2017 <- read_csv(here("raw_data/no2_by_grid_2017.csv"), skip = 6, 
+  col_names = c("uk_grid_code", "x", "y", "no2"))
+no2_2018 <- read_csv(here("raw_data/no2_by_grid_2018.csv"), skip = 6, 
+  col_names = c("uk_grid_code", "x", "y", "no2"))
+no2_2019 <- read_csv(here("raw_data/no2_by_grid_2019.csv"), skip = 6, 
+  col_names = c("uk_grid_code", "x", "y", "no2"))
+
+# Create empty data frame
+no2_all <- data_frame()
+
+# For each year, bind rows to one dataset
+for (i in 2010:2019) {
+  df_name <- paste0("no2_", i)
+  df_input <- as.name(df_name)
+
+  df <- eval(df_input) %>% 
+    mutate(year = i)
+
+no2_all <- bind_rows(no2_all, df)
+  }
+```

### 2021-09-14 documentation_and_analysis/draftbook.Rmd
Plot - EV vs Time in UK plot complete

@@
 ```{r}
+# Geospatial of EV Vehicles in the UK 2021 Q1
 uk_ev_map %>% 
@@
             position = "bottomright")
+```
+
+```{r}
+# Wrangling to create an EV count over time plot 
+uk_ev_longer <- uk_ev %>%
+  # Pivot longer to get year and count columns
+  pivot_longer(cols = c(x2021_q1:x2011_q4), names_to = c("year"), values_to = "no_of_ev") %>% 
+  # Filter so we only have UK as a whole data AND we only want final numbers of the year so Q4 
+  filter(region_local_authority_apr_2019_3 == "United Kingdom" & str_detect(year, "q4")) %>% 
+  # Simplify to just show year
+  mutate(year = case_when(str_detect(year, "2021") ~ "2021",
+         str_detect(year, "2020") ~ "2020",
+         str_detect(year, "2019") ~ "2019",
+         str_detect(year, "2018") ~ "2018",
+         str_detect(year, "2017") ~ "2017",
+         str_detect(year, "2016") ~ "2016",
+         str_detect(year, "2015") ~ "2015",
+         str_detect(year, "2014") ~ "2014",
+         str_detect(year, "2013") ~ "2013",
+         str_detect(year, "2012") ~ "2012",
+         str_detect(year, "2011") ~ "2011"),
+         year = as.numeric(year),
+         no_of_ev = as.numeric(no_of_ev))
+```
+
+```{r}
+uk_ev_longer %>% 
+  ggplot() +
+  aes(x = year, y = no_of_ev) +
+  geom_col(fill = "lawngreen") +
+  scale_x_continuous(breaks = c(2011:2020)) +
+  scale_y_continuous(breaks = seq(0, 220000, by = 20000), limits = c(0, 220000)) +
+  labs(title = "\nNumber of Electric Vehicles over time in the UK\n",
+       x = "\nYear\n",
+       y = "\nNumber of Electric Vehicles\n") +
+  theme_minimal() 
 ```

### 2021-09-14 documentation_and_analysis/draftbook.Rmd
Created heatmap of NO2 across UK

@@
 library(leaflet)
+library(leaflet.extras)
 ```
@@
 ```{r}
-air_pollution <- GET("https://uk-air.defra.gov.uk/data/sos/service?service=AQD&version=1.0.0&request=GetObservation&temporalFilter=om:phenomenonTime,After,yyyy-mm-ddT00:00:00.000Z") 
+air_pollution_cap <- GET("https://uk-air.defra.gov.uk/data/sos/service?service=AQD&request=GetCapabilities")
 ```
@@
 ```{r}
-air_pollution <- rbindlist(air_pollution$content, fill = TRUE) 
+air_pollution_cap <- rbindlist(air_pollution$content, fill = TRUE)
+```
+
+```{r}
+air_pollution_cap <- unnest(air_pollution_cap)
+```
+
+```{r}
+air_pollution <- GET("https://uk-air.defra.gov.uk/data/sos/service?service=SOS&version=2.0.0&request=GetObserved&observedProperty=http://dd.eionet.europa.eu/vocabulary/aq/pollutant/8") %>% 
+  content()
+```
+
+```{r}
+air_pollution$exceptions
+```
+
+```{r}
+air_pollution <- rbindlist(air_pollution$exceptions, fill = TRUE) 
 ```
@@
 ```
+
+```{r}
+# Remove missing NO2 grid values
+no2_clean <- no2_all %>% 
+  filter(no2 != "MISSING")
+```
+
+```{r}
+library(proj4)
+proj4string <- "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.9996012717 +x_0=400000 +y_0=-100000 +ellps=airy +datum=OSGB36 +units=m +no_defs"
+
+no2_clean_row <- no2_clean %>% 
+  rowid_to_column()
+
+# Source data
+xy <- no2_clean_row %>% 
+  select(x, y, rowid)
+
+# Transformed data
+pj <- project(xy, proj4string, inverse=TRUE)
+latlon <- data.frame(xy, lat=pj$y, lon=pj$x)
+final <-  merge(no2_clean_row, latlon, by.x = "rowid", by.y = "rowid") %>%
+  filter(year == 2019) %>% 
+  select(lat, lon, no2)
+
+```
+
+```{r}
+leaflet() %>%
+  addProviderTiles("CartoDB.Positron", options = providerTileOptions(noWrap = TRUE)) %>%
+  setView(lng = -4.2026, lat = 55.8, zoom = 4.7, options = list()) %>%
+  addHeatmap(data = final,
+             lng = ~lon, 
+             lat = ~lat, 
+             intensity = ~no2,
+             minOpacity = 0.05,
+             max = 5,
+             radius = 7)
+```

### 2021-09-16 documentation_and_analysis/draftbook.Rmd
Update draftbook.Rmd
Removal of unused data
Fixed heatmap (zoom issues)

@@
 
-# How many electric vehicles are on the road across the UK?
+# How many electric vehicles are on the road across the UK by LA?
 
@@
   left_join(uk_shape_file, by = c("ons_la_code_apr_2019" = "lad19cd")) %>% 
+  drop_na() %>% 
+  mutate(across(c(x2021_q1:x2011_q4), as.numeric)) %>% 
+  st_as_sf()
+```
+
+```{r}
+    pal <- colorBin("Greens", domain = uk_ev_map$x2021_q1, bins = c(0, 500, 1000, 2500, 5000, 10000, 15000))
+
+    uk_ev_map_labels <- sprintf(
+      "<strong>%s</strong><br/>%g Electric Vehicles",
+      uk_ev_map$region_local_authority_apr_2019_3, uk_ev_map$x2021_q1) %>% 
+      lapply(htmltools::HTML)
+```
+
+```{r}
+# Geospatial of EV Vehicles in the UK 2021 Q1
+uk_ev_map %>% 
+  leaflet() %>% 
+  setView(lng = -4.2026, lat = 55.8, zoom = 4.7, options = list()) %>%
+  addProviderTiles(providers$CartoDB.Positron) %>% 
+  addPolygons(fillColor = ~pal(x2021_q1),
+    weight = 0.1,
+    opacity = 0.9, 
+    color = "black",
+    fillOpacity = 0.8,
+    highlightOptions = highlightOptions(color = "green", weight = 2,
+                                        bringToFront = TRUE),
+    label = uk_ev_map_labels,
+    labelOptions = labelOptions(
+      style = list("font-weight" = "normal", padding = "3px 8px"),
+      textsize = "15px",
+      direction = "auto")) %>% 
+  addLegend(pal = pal, values = ~x2021_q1, opacity = 0.7, title = NULL,
+            position = "bottomright")
+```
+# How many electric vehicles are on the road across the UK by LA?
+
+```{r}
+# Reading in and skipping first 5 rows
+uk_ev_postcode <- read_ods(here("raw_data/ev_by_postcode.ods"), sheet = 2, skip = 6) %>% 
+  rename("postcode" = "Postcode District2")
+```
+
+```{r}
+# loading in shape file 
+uk_shape_file_postcode <- st_read(here("raw_data/postcode_shape/EX_Sample.shp")) %>%
+  clean_names() %>% 
+  st_simplify(dTolerance = 1000) %>%
+  st_transform("+proj=longlat +datum=WGS84") #%>% 
+#  select(lad19cd, long, lat, geometry) 
+```
+
+```{r}
+# Joining uk_ev + shape file 
+uk_ev_map_postcode <- uk_ev_postcode %>% 
+  left_join(uk_shape_file_postcode, by = c("ons_la_code_apr_2019" = "lad19cd")) %>% 
   drop_na() %>% 
@@
   select(lat, lon, no2) 
-
 ```
@@
 ```{r}
+final <- final %>% 
+  mutate(no2 = as.numeric(no2)) 
+```
+
+```{r}
+bins <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)
+pal <- colorBin("Spectral", domain = final$no2, bins = bins, na.color = "transparent", reverse = TRUE)
+
 leaflet() %>%
@@
              intensity = ~no2,
-             minOpacity = 0.05,
-             max = 5,
-             radius = 7)
+             minOpacity = 0.1,
+             max = 40,
+             radius = 1,
+             blur = 1) %>% 
+  addLegend(pal = pal, values = final$no2,
+                title="Average NO2 Conc")
 ```
+
+```{r}
+final %>% 
+  head()
+```
+
+```{r}
+write_csv(final, "no2_lat_long_data.csv")
+```
+
+```{r}
+no2_annual_mean <- read_ods(here("raw_data/NO2_tables.ods"), sheet = 3, skip = 2) %>% 
+  clean_names()
+```
+
+```{r}
+no2_annual_mean_all <- no2_annual_mean %>% 
+  filter(site == "All sites") %>% 
+  mutate(year = as.numeric(year))
+```
+
+```{r}
+no2_annual_mean_all %>% 
+  ggplot() +
+  aes(x = year, y = annual_mean_no2concentration_mg_m3) + 
+  geom_point() +
+  geom_smooth(se = FALSE) +
+  theme_minimal() +
+  scale_y_continuous(breaks = seq(0, 70, 10), limits = c(0, 70)) +
+  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
+  labs(x = "\nYear\n", 
+       y = "\nAnnual Mean NO2 Conc mg m3\n",
+       title = "\nNO2 over time in the UK\n") 
+```
+
+```{r}
+library(tidyverse)
+library(fable)
+library(tsibble)
+library(tsibbledata)
+library(ggfortify)
+```
+
+```{r}
+forecast_no2 <- no2_annual_mean_all %>% 
+  dplyr::select(year, annual_mean_no2concentration_mg_m3) %>% 
+  tsibble(index = year)
+```
+
+```{r}
+autoplot(forecast_no2)
+```
+
+```{r}
+fit <- forecast_no2 %>%
+  model(
+    mean_model = MEAN(annual_mean_no2concentration_mg_m3),
+    arima = ARIMA(annual_mean_no2concentration_mg_m3)
+  )
+fit
+```
+
+```{r}
+forecast_1 <- fit %>%
+  fabletools::forecast(h = 15)
+forecast_1
+```
+
+```{r}
+forecast_1 %>%
+  autoplot(forecast_no2) +
+  ggtitle("Forecasts for Australian beer sales") +
+  xlab("Year") +
+  guides(colour = guide_legend(title = "Forecast"))
+```

### 2021-09-17 documentation_and_analysis/draftbook.Rmd
NO2 Forecasts complete

@@
              minOpacity = 0.1,
-             max = 40,
+             max = 45,
              radius = 1,
@@
                 title="Average NO2 Conc")
-```
-
-```{r}
-final %>% 
-  head()
-```
-
-```{r}
-write_csv(final, "no2_lat_long_data.csv")
 ```
@@
   model(
-    mean_model = MEAN(annual_mean_no2concentration_mg_m3),
     arima = ARIMA(annual_mean_no2concentration_mg_m3)
@@
 forecast_1 <- fit %>%
-  fabletools::forecast(h = 15)
+  fabletools::forecast(h = "15 years")
 forecast_1
@@
 forecast_1 %>%
-  autoplot(forecast_no2) +
-  ggtitle("Forecasts for Australian beer sales") +
+  autoplot(forecast_no2, level = NULL) +
+  ggtitle("Forecasts for NO2 Conc over time") +
   xlab("Year") +
-  guides(colour = guide_legend(title = "Forecast"))
+  guides(colour = guide_legend(title = "Forecast")) +
+  scale_y_continuous(limit = c(0, 65))
 ```
+
+```{r}
+# Now set our training data from 1992 to 2006
+train <- forecast_no2 %>%
+  filter(year <= 2017)
+
+# run the model on the training set 
+fit_train <- train %>%
+  model(
+    arima = ARIMA(annual_mean_no2concentration_mg_m3)
+  )
+```
+
+```{r}
+# forecast from the training set
+forecast_test <- fit_train %>% 
+  fabletools::forecast(h = "4 years")
+
+# Plot forecasts against actual values
+forecast_test %>%
+  autoplot(train, level = NULL) +
+    autolayer(filter(forecast_no2, year >= 2017), color = "black") +
+    ggtitle("Forecasts for NO2 Conc over time") +
+    xlab("Year") + ylab("Megalitres") +
+    guides(colour=guide_legend(title="Forecast"))
+```
+
+```{r}
+# Binding 2014 and 2019
+no2_2014 <- read_csv(here("raw_data/no2_by_grid_2014.csv"), skip = 6, 
+  col_names = c("uk_grid_code", "x", "y", "no2")) %>% 
+  add_column(year = "2014")
+no2_2019 <- read_csv(here("raw_data/no2_by_grid_2019.csv"), skip = 6, 
+  col_names = c("uk_grid_code", "x", "y", "no2")) %>% 
+  add_column(year = "2019")
+
+no2_difference <- bind_rows(c(no2_2014, no2_2019))
+```
+
+```{r}
+# Diff in NO2 between 2014 & 2019
+no2_difference <- no2_difference %>% 
+  filter(no2 != "MISSING") %>% 
+  mutate(no2 = as.numeric(no2),
+         year = as.numeric(year)) %>% 
+  group_by(x, y) %>% 
+  pivot_wider(names_from = "year", values_from = "no2") %>% 
+  rename("x2014" = "2014",
+         "x2019" = "2019") %>% 
+  mutate(no2_diff_2014_2019 = x2014 - x2019) %>% 
+  ungroup()
+```
+
+```{r}
+# Converting NO2 diff from x y to lat long
+library(proj4)
+proj4string <- "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.9996012717 +x_0=400000 +y_0=-100000 +ellps=airy +datum=OSGB36 +units=m +no_defs"
+
+no2_diff_row <- no2_difference %>% 
+  rowid_to_column()
+
+# Source data
+xy <- no2_diff_row %>% 
+  select(x, y, rowid)
+
+# Transformed data
+pj <- project(xy, proj4string, inverse=TRUE)
+latlon <- data.frame(xy, lat=pj$y, lon=pj$x)
+no2_diff_final <-  merge(no2_diff_row, latlon, by.x = "rowid", by.y = "rowid") %>%
+  select(lat, lon, no2_diff_2014_2019) 
+```
+
+```{r}
+    pal <- colorBin("Greens", domain = no2_difference$no2_diff_2014_2019, bins = c(0, 500, 1000, 2500, 5000, 10000, 15000))
+
+    no2_diff_labels <- sprintf(
+      "<strong>%s</strong><br/>%g NO2 Diff",
+      no2_difference$uk_grid_code, no2_difference$no2_diff_2014_2019) %>% 
+      lapply(htmltools::HTML)
+```
+
+```{r}
+# Geospatial of EV Vehicles in the UK 2021 Q1
+no2_difference %>% 
+  leaflet() %>% 
+  setView(lng = -4.2026, lat = 55.8, zoom = 4.7, options = list()) %>%
+  addProviderTiles(providers$CartoDB.Positron) %>% 
+  addPolygons(fillColor = ~pal(no2_diff_2014_2019),
+    weight = 0.1,
+    opacity = 0.9, 
+    color = "black",
+    fillOpacity = 0.8,
+    highlightOptions = highlightOptions(color = "green", weight = 2,
+                                        bringToFront = TRUE),
+    label = no2_diff_labels,
+    labelOptions = labelOptions(
+      style = list("font-weight" = "normal", padding = "3px 8px"),
+      textsize = "15px",
+      direction = "auto")) %>% 
+  addLegend(pal = pal, values = ~no2_diff_2014_2019, opacity = 0.7, title = NULL,
+            position = "bottomright")
+```

### 2021-09-17 documentation_and_analysis/draftbook.Rmd
Pre draftbook tidy

@@
   mutate(no2_diff_2014_2019 = x2014 - x2019) %>% 
+#Remove this separation when put in cleaning script
+no2_difference <- no2_difference %>% 
   ungroup()
+#Remove this separation when put in cleaning script
+no2_difference <- no2_difference %>% 
+  drop_na()
+#Remove this separation when put in cleaning script
+# Changing all negative numbers to zero as these aren't of interest to us
+no2_difference <- no2_difference %>% 
+  mutate(no2_diff_2014_2019 = if_else(no2_diff_2014_2019 < 0, 0, no2_diff_2014_2019))
 ```
@@
 xy <- no2_diff_row %>% 
-  select(x, y, rowid)
+  dplyr::select(x, y, rowid)
 
@@
 no2_diff_final <-  merge(no2_diff_row, latlon, by.x = "rowid", by.y = "rowid") %>%
-  select(lat, lon, no2_diff_2014_2019) 
+  dplyr::select(lat, lon, no2_diff_2014_2019) 
 ```
@@
 ```{r}
-    pal <- colorBin("Greens", domain = no2_difference$no2_diff_2014_2019, bins = c(0, 500, 1000, 2500, 5000, 10000, 15000))
+# Geospatial of No2 diff 2014 to 2019
+bins <- c(0, 5, 10, 15, 20, 25, 30)
+pal <- colorBin("Spectral", domain = no2_diff_final$no2_diff_2014_2019, bins = bins, na.color = "transparent", reverse = TRUE)
 
-    no2_diff_labels <- sprintf(
-      "<strong>%s</strong><br/>%g NO2 Diff",
-      no2_difference$uk_grid_code, no2_difference$no2_diff_2014_2019) %>% 
-      lapply(htmltools::HTML)
+leaflet() %>%
+  addProviderTiles("CartoDB.Positron", options = providerTileOptions(noWrap = TRUE)) %>%
+  setView(lng = -4.2026, lat = 55.8, zoom = 4.7, options = list()) %>%
+  addHeatmap(data = no2_diff_final,
+             lng = ~lon,
+             lat = ~lat,
+             intensity = ~no2_diff_2014_2019,
+             minOpacity = 0.1,
+             max = 30,
+             radius = 1,
+             blur = 1) %>% 
+  addLegend(pal = pal, values = no2_diff_final$no2_diff_2014_2019,
+                title="Average NO2 Diff")
 ```
@@
 ```{r}
-# Geospatial of EV Vehicles in the UK 2021 Q1
 no2_difference %>% 
-  leaflet() %>% 
-  setView(lng = -4.2026, lat = 55.8, zoom = 4.7, options = list()) %>%
-  addProviderTiles(providers$CartoDB.Positron) %>% 
-  addPolygons(fillColor = ~pal(no2_diff_2014_2019),
-    weight = 0.1,
-    opacity = 0.9, 
-    color = "black",
-    fillOpacity = 0.8,
-    highlightOptions = highlightOptions(color = "green", weight = 2,
-                                        bringToFront = TRUE),
-    label = no2_diff_labels,
-    labelOptions = labelOptions(
-      style = list("font-weight" = "normal", padding = "3px 8px"),
-      textsize = "15px",
-      direction = "auto")) %>% 
-  addLegend(pal = pal, values = ~no2_diff_2014_2019, opacity = 0.7, title = NULL,
-            position = "bottomright")
+  ggplot() +
+  geom_histogram(aes(x = no2_diff_2014_2019))
 ```

### 2021-09-19 documentation_and_analysis/draftbook.Rmd
Charger geo spatial

@@
 ```{r}
-library(tidyverse)
 library(janitor)
@@
 library(leaflet.extras)
-```
-
-```{r}
-read_ods(here("raw_data/electric-vehicle-charging-device-statistics-june-2021.ods"))
-```
-
-```{r}
-getwd()
-```
-
-```{r}
-APE <- read_ods("raw_data/APE_site_data_tables.ods", sheet = 2)
-```
-
-```{r}
-NO2 <- read_ods("raw_data/NO2_tables.ods")
-```
-
-```{r}
-registry <- GET("http://chargepoints.dft.gov.uk/api/retrieve/registry/format/csv/") %>% 
-  content()
-```
-
-```{r}
-type <- GET("http://chargepoints.dft.gov.uk/api/retrieve/type/format/csv/") %>% 
-  content()
-```
-
-```{r}
-read_csv("Documents/GitHub/ev_climate_change_project/raw_data/no2_by_grid_2019.csv")
-```
-
-```{r}
-air_pollution_cap <- GET("https://uk-air.defra.gov.uk/data/sos/service?service=AQD&request=GetCapabilities")
-```
-
-```{r}
-air_pollution_cap <- rbindlist(air_pollution$content, fill = TRUE)
-```
-
-```{r}
-air_pollution_cap <- unnest(air_pollution_cap)
-```
-
-```{r}
-air_pollution <- GET("https://uk-air.defra.gov.uk/data/sos/service?service=SOS&version=2.0.0&request=GetObserved&observedProperty=http://dd.eionet.europa.eu/vocabulary/aq/pollutant/8") %>% 
-  content()
-```
-
-```{r}
-air_pollution$exceptions
-```
-
-```{r}
-air_pollution <- rbindlist(air_pollution$exceptions, fill = TRUE) 
-```
-
-```{r}
-air_pollution <- unnest(air_pollution)
-```
-
-# How many electric vehicles are on the road across the UK by LA?
-
-```{r}
-# Reading in and skipping first 5 rows
-uk_ev <- read_ods(here("raw_data/ev_by_la.ods"), sheet = 2, skip = 5)
-# Making row 1 the variable name
-names(uk_ev) <- uk_ev[1,] 
-# Removing row 1 
-uk_ev <- uk_ev[-1,] %>% 
-  clean_names()
+library(tidyverse)
 ```
@@
 # loading in shape file 
-uk_shape_file <- st_read(here("raw_data/Local_Authority_Districts__April_2019__UK_BFE_v2-shp/Local_Authority_Districts__April_2019__UK_BFE_v2.shp")) %>%
+uk_shape_file <- st_read(here("raw_data/LA_shape/Local_Authority_Districts__April_2019__UK_BFE_v2.shp")) %>%
   clean_names() %>% 
@@
   st_transform("+proj=longlat +datum=WGS84") %>% 
-  select(lad19cd, long, lat, geometry) 
+  dplyr::select(lad19cd, long, lat, geometry) 
 ```
@@
 ```{r}
+# Load in clean Electric Vehicles by Local Authority data
+uk_ev <- read_csv(here("clean_data/ev_by_la_clean.csv"))
 # Joining uk_ev + shape file 
@@
 ```{r}
+# Set colours and bins
 pal <- colorBin("Greens", domain = uk_ev_map$x2021_q1, bins = c(0, 500, 1000, 2500, 5000, 10000, 15000))
 
+# Set labels
 uk_ev_map_labels <- sprintf(
@@
              radius = 1,
-             blur = 1) %>% 
+             blur = 2) %>% 
   addLegend(pal = pal, values = no2_diff_final$no2_diff_2014_2019,
-                title="Average NO2 Diff")
+                title="Annual Mean NO2 Change between 2014 and 2019")
 ```
@@
 ```
+
+```{r}
+ev_charging <- read_csv(here("raw_data/electric-vehicle-charging-device-statistics-july-2021/EVCD_02-Table 1.csv"), skip = 6) %>% 
+  clean_names() %>% 
+  dplyr::select(year, month, total_devices, rapid_devices) %>% 
+  filter(rapid_devices != "NA") %>% 
+  mutate(total_devices = as.numeric(total_devices),
+         other_devices = total_devices - rapid_devices) %>% 
+  select(-total_devices) %>% 
+  pivot_longer(cols = c(other_devices, rapid_devices), names_to = "charger", values_to = "number_of_chargers") %>% 
+  filter(month == "October" | month == "July" & year == "2021") %>% 
+  mutate(charger = factor(charger, levels = c("rapid_devices", "other_devices")))
+```
+
+```{r}
+ev_charging %>% 
+  ggplot() +
+  aes(x = year, y = number_of_chargers, fill = charger) +
+  geom_col() +
+  scale_y_continuous(breaks = seq(0, 30000, 2500)) +
+  coord_flip() +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 270, vjust = 0.5, hjust=1)) +
+  scale_fill_manual(labels = c("Rapid Charger", "Standard Charger"), values = c("#e41a1c", "#4daf4a")) +
+  labs(title = "\nNumber of Electric Vehicle Chargers Growth in the UK\n",
+       x = "\nYear\n",
+       y = "\nNumber of Chargers\n",
+       fill = "Charger Type")
+```
+
+```{r}
+ev_charging_rapid <- read_csv(here("raw_data/electric-vehicle-charging-device-statistics-july-2021/EVCD_01b-Table 1.csv"), skip = 6) %>% 
+  clean_names() %>% 
+  select(la_region_code, local_authority_region_name, jul_21) %>% 
+  rename("count_rapid" = jul_21) 
+ev_charging_total <- read_csv(here("raw_data/electric-vehicle-charging-device-statistics-july-2021/EVCD_01a-Table 1.csv"), skip = 6) %>% 
+  clean_names() %>% 
+  select(la_region_code, local_authority_region_name, jul_21) %>% 
+  rename("count_total" = jul_21)
+
+ev_charging_geo <- inner_join(ev_charging_rapid, ev_charging_total, by = "la_region_code") %>% 
+  drop_na() %>% 
+  select(la_region_code, local_authority_region_name.x, count_rapid, count_total) %>% 
+  filter(str_detect(local_authority_region_name.x, "[a-z][a-z]")#,
+         #count_rapid != "-") %>% 
+  mutate(local_authority_region_name.x = str_remove_all(local_authority_region_name.x, "[(](Met County)[)]"))
+```
+
+```{r}
+# Joining ev_charging_geo + shape file 
+ev_charging_map <- ev_charging_geo %>% 
+  left_join(uk_shape_file, by = c("la_region_code" = "lad19cd")) %>% 
+  drop_na() %>% 
+  mutate(across(c(count_rapid:count_total), as.numeric)) %>% 
+  st_as_sf()
+```
+
+```{r}
+# Set colours and bins
+pal <- colorBin("Reds", domain = ev_charging_map$count_total, bins = c(0, 50, 100, 250, 500, 750, 1000, 1250))
+
+# Set labels
+ev_charging_map_labels <- sprintf(
+  "<strong>%s</strong><br/>%g Chargers",
+  ev_charging_map$local_authority_region_name.x, ev_charging_map$count_total, ev_charging_map$count_rapid) %>% 
+  lapply(htmltools::HTML)
+```
+
+```{r}
+# Geospatial of EV Vehicles in the UK 2021 Q1
+ev_charging_map %>% 
+  leaflet() %>% 
+  setView(lng = -4.2026, lat = 55.8, zoom = 4.7, options = list()) %>%
+  addProviderTiles(providers$CartoDB.Positron) %>% 
+  addPolygons(fillColor = ~pal(count_total),
+    weight = 0.1,
+    opacity = 0.9, 
+    color = "black",
+    fillOpacity = 0.8,
+    highlightOptions = highlightOptions(color = "green", weight = 2,
+                                        bringToFront = TRUE),
+    label = ev_charging_map_labels,
+    labelOptions = labelOptions(
+      style = list("font-weight" = "normal", padding = "3px 8px"),
+      textsize = "15px",
+      direction = "auto")) %>% 
+  addLegend(pal = pal, values = ~count_total, opacity = 0.7, title = NULL,
+            position = "bottomright")
+```
+
+```{r}
+ev_charging_map %>% 
+  filter(count_total > 500)
+```

### 2021-09-19 data_cleaning_scripts/ev_by_la_cleaning_script.R

@@ -1,0 +1,14 @@
+library(tidyverse)
+library(janitor)
+library(here)
+library(readODS)
+  
+# Reading in and skipping first 5 rows
+uk_ev <- read_ods(here("raw_data/ev_by_la.ods"), sheet = 2, skip = 5)
+# Making row 1 the variable name
+names(uk_ev) <- uk_ev[1,] 
+# Removing row 1 
+uk_ev <- uk_ev[-1,] %>% 
+  clean_names()
+
+write_csv(uk_ev, "clean_data/ev_by_la_clean.csv")

### 2021-09-19 documentation_and_analysis/draftbook.Rmd
Rapid + Total Charger Geospatial Completed

@@
   clean_names() %>% 
-  select(la_region_code, local_authority_region_name, jul_21) %>% 
-  rename("count_rapid" = jul_21) 
+  select(la_region_code, local_authority_region_name, apr_21) %>% 
+  rename("count_rapid" = apr_21)
 ev_charging_total <- read_csv(here("raw_data/electric-vehicle-charging-device-statistics-july-2021/EVCD_01a-Table 1.csv"), skip = 6) %>% 
   clean_names() %>% 
-  select(la_region_code, local_authority_region_name, jul_21) %>% 
-  rename("count_total" = jul_21)
+  select(la_region_code, local_authority_region_name, apr_21) %>% 
+  rename("count_total" = apr_21)
 
@@
   select(la_region_code, local_authority_region_name.x, count_rapid, count_total) %>% 
-  filter(str_detect(local_authority_region_name.x, "[a-z][a-z]")#,
-         #count_rapid != "-") %>% 
-  mutate(local_authority_region_name.x = str_remove_all(local_authority_region_name.x, "[(](Met County)[)]"))
+  filter(str_detect(local_authority_region_name.x, "[a-z][a-z]"), 
+         count_rapid != "-") %>% 
+  mutate(local_authority_region_name.x = str_remove_all(local_authority_region_name.x, "[(](Met County)[)]"),
+         local_authority_region_name.x = str_remove_all(local_authority_region_name.x, "[(]abolished \\w+ \\d+[)]"))
 ```
@@
 ev_charging_map <- ev_charging_geo %>% 
-  left_join(uk_shape_file, by = c("la_region_code" = "lad19cd")) %>% 
-  drop_na() %>% 
+  inner_join(uk_shape_file, by = c("la_region_code" = "lad19cd")) %>% 
   mutate(across(c(count_rapid:count_total), as.numeric)) %>% 
@@
 # Set colours and bins
-pal <- colorBin("Reds", domain = ev_charging_map$count_total, bins = c(0, 50, 100, 250, 500, 750, 1000, 1250))
+pal <- colorBin("Reds", domain = ev_charging_map$count_total, bins = 10)
 
@@
 ```{r}
-# Geospatial of EV Vehicles in the UK 2021 Q1
+# Geospatial of EV Total Chargers in the UK April 21
 ev_charging_map %>% 
@@
 ```{r}
+# Set colours and bins
+pal <- colorBin("Reds", domain = ev_charging_map$count_rapid, bins = 10)
+
+# Set labels
+ev_charging_map_labels <- sprintf(
+  "<strong>%s</strong><br/>%g Chargers",
+  ev_charging_map$local_authority_region_name.x, ev_charging_map$count_rapid, ev_charging_map$count_total) %>% 
+  lapply(htmltools::HTML)
+```
+
+```{r}
+# Geospatial of EV Rapid Chargers in the UK April 21
 ev_charging_map %>% 
-  filter(count_total > 500)
+  leaflet() %>% 
+  setView(lng = -4.2026, lat = 55.8, zoom = 4.7, options = list()) %>%
+  addProviderTiles(providers$CartoDB.Positron) %>% 
+  addPolygons(fillColor = ~pal(count_rapid),
+    weight = 0.1,
+    opacity = 0.9, 
+    color = "black",
+    fillOpacity = 0.8,
+    highlightOptions = highlightOptions(color = "green", weight = 2,
+                                        bringToFront = TRUE),
+    label = ev_charging_map_labels,
+    labelOptions = labelOptions(
+      style = list("font-weight" = "normal", padding = "3px 8px"),
+      textsize = "15px",
+      direction = "auto")) %>% 
+  addLegend(pal = pal, values = ~count_rapid, opacity = 0.7, title = NULL,
+            position = "bottomright")
 ```

### 2021-09-20 documentation_and_analysis/draftbook.Rmd
New car reg vs fuel type

@@
 
-no2_difference <- bind_rows(c(no2_2014, no2_2019))
+no2_difference <- bind_rows(no2_2014, no2_2019)
 ```
@@
 #Remove this separation when put in cleaning script
-no2_difference <- no2_difference %>% 
-  ungroup()
+  ungroup() %>% 
 #Remove this separation when put in cleaning script
-no2_difference <- no2_difference %>% 
-  drop_na()
+  drop_na() %>% 
 #Remove this separation when put in cleaning script
 # Changing all negative numbers to zero as these aren't of interest to us
-no2_difference <- no2_difference %>% 
   mutate(no2_diff_2014_2019 = if_else(no2_diff_2014_2019 < 0, 0, no2_diff_2014_2019))
@@
   clean_names() %>% 
-  select(la_region_code, local_authority_region_name, apr_21) %>% 
-  rename("count_rapid" = apr_21)
+  select(la_region_code, local_authority_region_name, x6) %>% 
+  rename("count_rapid" = x6)
 ev_charging_total <- read_csv(here("raw_data/electric-vehicle-charging-device-statistics-july-2021/EVCD_01a-Table 1.csv"), skip = 6) %>% 
   clean_names() %>% 
-  select(la_region_code, local_authority_region_name, apr_21) %>% 
-  rename("count_total" = apr_21)
+  select(la_region_code, local_authority_region_name, x6) %>% 
+  rename("count_total" = x6)
 
@@
 ev_charging_map_labels <- sprintf(
-  "<strong>%s</strong><br/>%g Chargers",
+  "<strong>%s</strong><br/>%g Chargers per 100k Population",
   ev_charging_map$local_authority_region_name.x, ev_charging_map$count_total, ev_charging_map$count_rapid) %>% 
@@
 ```{r}
-# Geospatial of EV Total Chargers in the UK April 21
+# Geospatial of EV Chargers Per 100k people in the UK April 21
 ev_charging_map %>% 
@@
 ```{r}
-# Geospatial of EV Rapid Chargers in the UK April 21
+# Geospatial of EV Rapid Chargers per 100k people in the UK April 21
 ev_charging_map %>% 
@@
 ```
+
+```{r}
+uk_ev_map_2021 <- uk_ev_map %>% 
+  select(ons_la_code_apr_2019, region_local_authority_apr_2019_3, x2021_q1, geometry)
+
+# filter for desired polygon
+highlands <- uk_ev_map_2021 %>% 
+  filter(region_local_authority_apr_2019_3 == "Highland")
+
+# st_join seems less dirty
+ev_no2 <- final %>% 
+  st_as_sf(coords = c("lon", "lat"), crs = st_crs(highlands)) %>% 
+  st_join(uk_ev_map_2021, join = st_intersects, left = FALSE) %>% 
+  group_by(region_local_authority_apr_2019_3) %>% 
+  mutate(mean_no2 = sum(no2))
+```
+
+```{r}
+ev_no2 %>% 
+  ggplot() +
+  aes(x = x2021_q1, y = mean_no2) +
+  geom_point() +
+  geom_smooth(se = TRUE)
+```
+
+```{r}
+population <- read_csv(here("raw_data/ukpopestimatesmid2020on2021geography/MYE2 - Persons-Table 1.csv"), skip = 7) %>% 
+  clean_names() %>% 
+  select(code, name, all_ages) %>% 
+  mutate(pop_per_100k = all_ages/100000)
+```
+
+```{r}
+ev_no2_tbl <- tibble(ev_no2) %>% 
+  select(-c(no2, geometry)) %>% 
+  unique()
+```
+
+```{r}
+ev_pop_100k <- ev_no2_tbl %>% 
+  left_join(population, by = c("ons_la_code_apr_2019" = "code")) %>% 
+  mutate(ev_per_100k = x2021_q1/pop_per_100k) 
+```
+
+```{r}
+ev_pop_100k %>% 
+  ggplot() +
+  aes(x = ev_per_100k, y = mean_no2) +
+  geom_point() +
+  geom_smooth(se = TRUE)
+```
+
+```{r}
+library(cluster)
+library(factoextra)
+```
+
+```{r}
+ev_scale <- ev_pop_100k %>% 
+  select(region_local_authority_apr_2019_3, x2021_q1, mean_no2) %>% 
+  mutate_if(is.numeric, scale)
+```
+
+```{r}
+ev_pop_100k_scale <- ev_pop_100k %>% 
+  select(region_local_authority_apr_2019_3, ev_per_100k, mean_no2) %>% 
+  mutate_if(is.numeric, scale)
+```
+
+```{r}
+ev_scale %>% 
+  ggplot() +
+  aes(x = x2021_q1, y = mean_no2) +
+  geom_point() +
+  geom_smooth(method = lm)
+```
+
+```{r}
+ev_pop_100k_scale %>% 
+  ggplot() +
+  aes(x = ev_per_100k, y = mean_no2) +
+  geom_point() +
+  geom_smooth(method = lm)
+```
+
+```{r}
+library(corrplot)
+```
+```{r}
+ev_pop_100k_scale %>% 
+  as_tibble() %>%
+  pivot_longer(-region_local_authority_apr_2019_3,
+               names_to = "type", 
+               values_to = "value") %>% 
+  drop_na() %>% 
+  group_by(type) %>%
+  summarise(mean = round(mean(value)), 
+            sd = sd(value))
+```
+```{r}
+ev_pop_100k_scale_numeric <- ev_pop_100k_scale %>% 
+  select(-region_local_authority_apr_2019_3)
+```
+
+```{r}
+corrplot(cor(ev_pop_100k_scale_numeric), method = "number", type = "lower")
+```
+
+```{r}
+car_by_fuel_gb <- read_ods(here("raw_data/car_by_fuel_by_year_ALL_UK.ods"), skip = 7) %>% 
+  clean_names() %>% 
+  head(58) %>% 
+  mutate(across(c(petrol:zero_emission8), as.numeric)) %>% 
+  filter(petrol >= 100) %>% 
+  mutate(traditional_fuel = total - alternative_fuels7 - zero_emission8) %>% 
+  select(c("year", "petrol", "diesel", "alternative_fuels7", "zero_emission8")) %>% 
+  pivot_longer(cols = c(petrol:zero_emission8), names_to = c("fuel_type"), values_to = c("count"))
+```
+
+```{r}
+car_by_fuel_gb %>% 
+  ggplot() +
+  aes(x = year, y = count, colour = fuel_type) +
+  geom_smooth(method = lm) +
+  geom_point() 
+```
+
+```{r}
+new_car_by_fuel_gb <- read_ods(here("raw_data/new_car_by_fuel.ods"), skip = 7) %>% 
+  clean_names() %>% 
+  head(21) %>% 
+  mutate(across(c(petrol:zero_emission8), as.numeric)) %>% 
+  filter(petrol >= 100) %>% 
+  mutate(traditional_fuel = total - alternative_fuels7 - zero_emission8) %>% 
+  select(c("date", "petrol", "diesel", "alternative_fuels7", "zero_emission8")) %>% 
+  pivot_longer(cols = c(petrol:zero_emission8), names_to = c("fuel_type"), values_to = c("count"))
+```
+
+```{r}
+new_car_by_fuel_gb %>% 
+  ggplot() +
+  aes(x = date, y = count, colour = fuel_type) +
+  geom_line(aes(group = fuel_type)) +
+  theme_minimal() +
+  labs(title = "\nNew Car Registrations by Fuel Type\n",
+       x = "\nYear\n",
+       y = "\nNumber of Cars ('000s)\n",
+       colour = "Fuel Type") +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  scale_colour_manual(labels = c("Alternative Fuels", "Diesel", "Petrol", "Zero Emission"), values = c("#abdda4", "#d7191c", "#fdae61", "#2b83ba")) 
+```

### 2021-09-21 documentation_and_analysis/draftbook.Rmd
log ev_no2 graph complete

@@
 library(tidyverse)
+library(grDevices)
 ```
@@
             position = "bottomright")
-```
-# How many electric vehicles are on the road across the UK by LA?
-
-```{r}
-# Reading in and skipping first 5 rows
-uk_ev_postcode <- read_ods(here("raw_data/ev_by_postcode.ods"), sheet = 2, skip = 6) %>% 
-  rename("postcode" = "Postcode District2")
-```
-
-```{r}
-# loading in shape file 
-uk_shape_file_postcode <- st_read(here("raw_data/postcode_shape/EX_Sample.shp")) %>%
-  clean_names() %>% 
-  st_simplify(dTolerance = 1000) %>%
-  st_transform("+proj=longlat +datum=WGS84") #%>% 
-#  select(lad19cd, long, lat, geometry) 
-```
-
-```{r}
-# Joining uk_ev + shape file 
-uk_ev_map_postcode <- uk_ev_postcode %>% 
-  left_join(uk_shape_file_postcode, by = c("ons_la_code_apr_2019" = "lad19cd")) %>% 
-  drop_na() %>% 
-  mutate(across(c(x2021_q1:x2011_q4), as.numeric)) %>% 
-  st_as_sf()
 ```
@@
   aes(x = year, y = no_of_ev) +
-  geom_col(fill = "lawngreen") +
+  geom_col(fill = "#08a44c") +
   scale_x_continuous(breaks = c(2011:2020)) +
@@
 ```{r}
-bins <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)
+bins <- 10
 pal <- colorBin("Spectral", domain = final$no2, bins = bins, na.color = "transparent", reverse = TRUE)
@@
   addLegend(pal = pal, values = final$no2,
-                title="Average NO2 Conc")
+                title="Average NO2 Conc in 2019")
 ```
@@
   aes(x = year, y = annual_mean_no2concentration_mg_m3) + 
-  geom_point() +
-  geom_smooth(se = FALSE) +
+  geom_line() +
   theme_minimal() +
-  scale_y_continuous(breaks = seq(0, 70, 10), limits = c(0, 70)) +
-  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
+  scale_y_continuous(breaks = seq(0, 65, 5), limits = c(22, 62)) +
+  scale_x_continuous(breaks = seq(1997, 2020, 1)) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
   labs(x = "\nYear\n", 
        y = "\nAnnual Mean NO2 Conc mg m3\n",
-       title = "\nNO2 over time in the UK\n") 
+       title = "\nNO2 over time in the UK\n") +
+png("no2_time_uk.png", units="in", width=8, height=5, res=300)
+# insert ggplot code
+dev.off()
 ```
@@
   guides(colour = guide_legend(title = "Forecast")) +
-  scale_y_continuous(limit = c(0, 65))
+    theme_minimal() +
+  scale_y_continuous(breaks = seq(0, 65, 5), limits = c(0, 62)) +
+  scale_x_continuous(breaks = seq(1997, 2034, 3)) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  labs(x = "\nYear\n", 
+       y = "\nAnnual Mean NO2 Conc mg m3\n",
+       title = "\nNO2 over time in the UK\n") +
+  scale_y_continuous(limit = c(0, 65)) +
+  png("no2_time_forecast.png", units="in", width=8, height=5, res=300)
+# insert ggplot code
+dev.off()
 ```
@@
        y = "\nNumber of Chargers\n",
-       fill = "Charger Type")
+       fill = "Charger Type") +
+  png("no_chargers_time_uk.png", units="in", width=8, height=5, res=300)
+
 ```
@@
   left_join(population, by = c("ons_la_code_apr_2019" = "code")) %>% 
-  mutate(ev_per_100k = x2021_q1/pop_per_100k) 
+  mutate(ev_per_100k = x2021_q1/pop_per_100k) %>% 
+  drop_na()
 ```
@@
 ```{r}
-ev_pop_100k %>% 
+ev_pop_100k_log <- ev_pop_100k %>% 
+  mutate(ev_per_100k = log(ev_per_100k))
+```
+
+```{r}
+ev_pop_100k_log %>% 
   ggplot() +
@@
   geom_point() +
-  geom_smooth(se = TRUE)
+  geom_smooth(se = TRUE, method = lm)
 ```
@@
 ```{r}
-ev_scale <- ev_pop_100k %>% 
+ev_scale <- ev_pop_100k_log %>% 
   select(region_local_authority_apr_2019_3, x2021_q1, mean_no2) %>% 
@@
 ```{r}
-ev_pop_100k_scale <- ev_pop_100k %>% 
+ev_pop_100k_scale_log <- ev_pop_100k_log %>% 
   select(region_local_authority_apr_2019_3, ev_per_100k, mean_no2) %>% 
@@
 ```{r}
-ev_scale %>% 
+ev_scale_log %>% 
   ggplot() +
@@
 ```{r}
-ev_pop_100k_scale %>% 
+ev_pop_100k_scale_log %>% 
   ggplot() +
@@
   select(c("date", "petrol", "diesel", "alternative_fuels7", "zero_emission8")) %>% 
-  pivot_longer(cols = c(petrol:zero_emission8), names_to = c("fuel_type"), values_to = c("count"))
+  pivot_longer(cols = c(petrol:zero_emission8), names_to = c("fuel_type"), values_to = c("count")) %>% 
+  mutate(date = as.numeric(date))
 ```
@@
   aes(x = date, y = count, colour = fuel_type) +
-  geom_line(aes(group = fuel_type)) +
+  geom_line(aes(group = fuel_type), size = 1.25) +
   theme_minimal() +
+  scale_y_continuous(breaks = seq(0, 2250, 250)) +
+  scale_x_continuous(limits = c(2001, 2021), breaks = seq(2001, 2020, 1)) + 
   labs(title = "\nNew Car Registrations by Fuel Type\n",
@@
        colour = "Fuel Type") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  scale_colour_manual(labels = c("Alternative Fuels", "Diesel", "Petrol", "Zero Emission"), values = c("#abdda4", "#d7191c", "#fdae61", "#2b83ba")) 
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+  panel.grid.minor = element_blank()) +
+  scale_colour_manual(labels = c("Alternative Fuels", "Diesel", "Petrol", "Zero Emission"), values = c("#abdda4", "#d7191c", "#fdae61", "#2b83ba")) +
+  theme(legend.position = "none")  +
+  annotate("text", x = 2006, y = 1750, label = "Petrol", colour = "#fdae61", fontface = 2) +
+  annotate("text", x = 2005, y = 1100, label = "Diesel", colour = "#d7191c", fontface = 2) +
+  annotate("text", x = 2016, y = 300, label = "Alternative Fuels", colour = "#abdda4", fontface = 2) +
+  annotate("text", x = 2008, y = 150, label = "Zero Emissions", colour = "#2b83ba", fontface = 2) +
+annotate("text", x = 2020.60, y = 950, label = "987", colour = "#fdae61", fontface = 2) +
+  annotate("text", x = 2020.60, y = 250, label = "295", colour = "#d7191c", fontface = 2) +
+  annotate("text", x = 2020.60, y = 450, label = "338", colour = "#abdda4", fontface = 2) +
+  annotate("text", x = 2020.60, y = 50, label = "106", colour = "#2b83ba", fontface = 2) + 
+png("test.png", units="in", width=8, height=5, res=300)
+# insert ggplot code
+dev.off()
 ```
+
+```{r}
+uk_ev_top <- uk_ev %>% 
+  mutate(across(c(x2021_q1:x2011_q4), as.numeric)) %>% 
+  filter(region_local_authority_apr_2019_3 == "Milton Keynes" |
+         region_local_authority_apr_2019_3 == "Swindon" |
+         region_local_authority_apr_2019_3 == "Slough" | 
+         region_local_authority_apr_2019_3 == "Leeds") %>% 
+  pivot_longer(cols = c(x2021_q1:x2011_q4), names_to = c("year"), values_to = "no_of_ev") %>% 
+  filter(str_detect(year, "q4")) %>% 
+  mutate(year = str_sub(year, 2, -4))
+```
+
+```{r}
+uk_ev_top <- uk_ev_top %>% 
+  left_join(uk_shape_file, by = c("ons_la_code_apr_2019" = "lad19cd")) %>% 
+  drop_na() %>% 
+  st_as_sf()
+```
+
+```{r}
+# Converting NO2 diff from x y to lat long
+library(proj4)
+proj4string <- "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.9996012717 +x_0=400000 +y_0=-100000 +ellps=airy +datum=OSGB36 +units=m +no_defs"
+
+no2_all_clean <- no2_clean %>% 
+  rowid_to_column()
+
+# Source data
+xy <- no2_all_clean %>% 
+  dplyr::select(x, y, rowid)
+
+# Transformed data
+pj <- project(xy, proj4string, inverse=TRUE)
+latlon <- data.frame(xy, lat=pj$y, lon=pj$x)
+no2_all_final <-  merge(no2_all_clean, latlon, by.x = "rowid", by.y = "rowid") %>%
+  dplyr::select(lat, lon, no2, year) 
+```
+
+```{r}
+uk_ev_top_2021 <- uk_ev_top %>% 
+  select(ons_la_code_apr_2019, region_local_authority_apr_2019_3, year, no_of_ev)
+
+# filter for desired polygon
+#highlands <- uk_ev_stockport %>% 
+ # filter(region_local_authority_apr_2019_3 == "Highland")
+
+# st_join seems less dirty
+ev_no2_all_years <- no2_all_final %>% 
+  st_as_sf(coords = c("lon", "lat"), crs = st_crs(uk_ev_top_2021)) %>%
+  st_join(uk_ev_top_2021, join = st_intersects, left = FALSE) 
+```
+
+```{r}
+ev_no2_all_year_top <- tibble(ev_no2_all_years) %>% 
+  filter(year.x == year.y) %>% 
+  mutate(no2 = as.numeric(no2)) %>% 
+  group_by(region_local_authority_apr_2019_3, year.x) %>% 
+  mutate(mean_no2 = mean(no2))
+```
+
+```{r}
+ev_no2_all_year_top %>% 
+  ggplot() +
+  geom_line(aes(x = year.x, y = no_of_ev, colour = region_local_authority_apr_2019_3)) +
+  scale_x_continuous(breaks = c(2011:2019)) +
+  scale_y_continuous(breaks = seq(0, 2750, 250)) +
+  labs(title = "High EV Local Authority over Years",
+       x = "Year",
+       y = "Number of Electric Vehicles",
+       colour = "Local Authority") +
+  theme_minimal()
+```
+
+```{r}
+ev_no2_all_year_top %>% 
+  ggplot() +
+  geom_line(aes(x = year.x, y = mean_no2, colour = region_local_authority_apr_2019_3)) +
+  scale_x_continuous(breaks = c(2011:2019)) +
+  scale_y_continuous(breaks = c(10:28)) +
+  labs(title = "Mean NO2 per Year by Local Authority",
+       x = "Year",
+       y = "Mean NO2",
+       colour = "Local Authority") +
+  theme_minimal()
+```

### 2021-09-21 documentation_and_analysis/draftbook.Rmd
Presentation complete

@@
 ```{r}
-no2_difference %>% 
+no2_difference_log <- no2_difference %>% 
+  mutate(no2_diff_2014_2019_log = log(no2_diff_2014_2019))
+```
+
+```{r}
+no2_difference_log %>% 
   ggplot() +
-  geom_histogram(aes(x = no2_diff_2014_2019))
+  geom_histogram(aes(x = no2_diff_2014_2019_log))
 ```
@@
 ev_pop_100k_log <- ev_pop_100k %>% 
-  mutate(ev_per_100k = log(ev_per_100k))
+  mutate(ev_per_100k_log = log(ev_per_100k),
+         mean_no2_log = log(mean_no2))
 ```
@@
   ggplot() +
-  aes(x = ev_per_100k, y = mean_no2) +
+  aes(x = ev_per_100k_log, y = mean_no2_log) +
   geom_point() +
-  geom_smooth(se = TRUE, method = lm)
+  geom_smooth(se = FALSE, method = lm, colour = "#08a44c") +
+  theme_minimal() +
+  labs(title = "\nElectric Vehicles vs Mean NO2 Conc for each UK Local Authority in 2019\n",
+       x = "\nElectric Vehicle per 100k population (log)\n",
+       y = "\nMean NO2 Conc (log)\n") +
+  png("ev_no2_log.png", units="in", width=8, height=5, res=300)
 ```
@@
 ```
+
+```{r}
+ev_charging_map_2021 <- ev_charging_map %>% 
+  select(la_region_code, local_authority_region_name.x, count_rapid, count_total, geometry)
+
+# st_join seems less dirty
+ev_charging_no2 <- final %>% 
+  st_as_sf(coords = c("lon", "lat"), crs = st_crs(ev_charging_map_2021)) %>% 
+  st_join(ev_charging_map_2021, join = st_intersects, left = FALSE) %>% 
+  group_by(local_authority_region_name.x) %>% 
+  mutate(mean_no2 = mean(no2))
+```
+
+```{r}
+ev_charging_no2_log <- ev_charging_no2 %>% 
+  mutate(count_rapid_log = log(count_rapid),
+         count_total_log = log(count_total),
+         mean_no2_log = log(mean_no2))
+```
+
+```{r}
+ev_charging_no2 %>% 
+  ggplot() +
+  aes(x = count_rapid, y = mean_no2) +
+  geom_point() +
+  geom_smooth(se = TRUE, method = lm)
+```
+
+```{r}
+ev_charging_no2 %>% 
+  ggplot() +
+  aes(x = count_total, y = mean_no2) +
+  geom_point() +
+  geom_smooth(se = TRUE, method = lm)
+```
+
+```{r}
+ev_charging_no2_log %>% 
+  ggplot() +
+  aes(x = count_rapid_log, y = mean_no2_log) +
+  geom_point() +
+  geom_smooth(se = TRUE, method = lm)
+```
+
+```{r}
+ev_charging_no2_log %>% 
+  ggplot() +
+  aes(x = count_total_log, y = mean_no2_log) +
+  geom_point() +
+  geom_smooth(se = TRUE, method = lm, colour = "#08a44c") +
+  theme_minimal() +
+  labs(title = "\nEV Chargers vs Mean NO2 Conc for each UK Local Authority in 2019\n",
+       x = "\nTotal Number of EV Chargers (log)\n",
+       y = "\nMean NO2 Conc (log)\n") +
+  png("ev_charger_no2_log.png", units="in", width=8, height=5, res=300)
+```
+
+```{r}
+income_england <- read_csv(here("raw_data/incomeestimatesforsmallareasdatasetfinancialyearending20181 2/Total annual income-Table 1.csv"), skip = 4) %>% 
+  clean_names() %>% 
+  select(local_authority_code, local_authority_name, total_annual_income) %>% 
+  group_by(local_authority_code) %>% 
+  mutate(mean_total_annual_income =  mean(total_annual_income)) %>% 
+  select(-total_annual_income) %>% 
+  unique()
+```
+
+```{r}
+uk_ev_income <- uk_ev %>% 
+  inner_join(income_england, by = c("ons_la_code_apr_2019" = "local_authority_code")) %>% 
+  select(region_local_authority_apr_2019_3, x2021_q1, mean_total_annual_income) %>% 
+  mutate(x2021_q1 = as.numeric(x2021_q1),
+         x2021_q1_log = log(x2021_q1))
+```
+
+```{r}
+uk_ev_income %>% 
+  ggplot() +
+  aes(x = x2021_q1_log, y = mean_total_annual_income) +
+  geom_point() +
+  geom_smooth(method = lm, colour = "#08a44c", se = FALSE) +
+  theme_minimal() +
+  labs(title = "\nEVs vs Mean Annual Income for Local Authorities in England and Wales\n",
+       x = "\nTotal Number of EV Chargers (log)\n",
+       y = "\nMean Total Annual Income\n") +
+  png("ev_income.png", units="in", width=8, height=5, res=300)
+```

### 2021-09-23 documentation_and_analysis/the_road_to_sustinability_documentation.Rmd
Documentation complete

---
title: "The Road to Sustainability - Final Project"
output:
  html_document:
    toc: yes
    toc_float: yes
    df_print: paged
    pdf_document: default
  word_document:
    toc: yes
---

## Context

#### Business intelligence and data-driven decision making

The Department of Transport and the Department for Environment, Food and Rural Affairs will gain an understanding of trends to date and a forward look at how they should deliver a financial package towards their vehicular zero emission goals. They will be able to make better decisions as they will have a more rounded view of where the UK currently stands and have an idea of how to measure impact going forward.

#### Domain knowledge and the business context

The Department for Transport and the Department for Environment, Food and Rural Affairs are responsible for transport and the environment respectively. Together they have a combined target to reduce carbon emissions on UK roads to battle climate change. My analysis fits this aim as it shows them insights into data they already hold and gives them a plan of where to allocate resources to tackle this issue.

## Data

#### Internal and external data sources

Internal

- Department for Transport
- Department for Environment, Food and Rural Affairs 
- Office for National Statistics

External

- None

#### Types of data

Numerical Data

- Continuous - Number of electric vehichles, NO2 concentration, etc
- Discrete - None

Categorical Data

- Nominal - Fuel type, charger type, etc
- Ordinal - Year, etc

#### Data formats

My data mainly came in ODS format. I had a few in CSV and a single Shape file.

#### Data quality and bias

This is a complete publically available dataset which is collected from official vehicle registers and air sensors. For that reason, there is no obvious bias in the data.

## Ethics

#### Ethical issues in data sourcing and extraction

The data that was sourced from vehicle registers will not have consent from the vehicle owners. This could be a small ethical concern over usage consent.

#### Ethical implications of business requirements

Some people may oppose moving away from traditional fuel vehicles towards zero emission vehicles and so we are using their vehicle data to study a business requirement which is against their beliefs.

## Analysis

#### Stages in the data analysis process

- Planning & Trello board build
- Git repo creation and management
- Choosing datasets
- Data cleaning
- Data wrangling
- Deciding on visualisations
- Creating visualisations
- Gaining feedback from experts
- Presentation building
- Creating documentation

#### Tools for data analysis

- Trello (planning & task allocation)
- Git/GitHub (collaboration & version control)
- R Studio (cleaning data and creating visualisations)

#### Descriptive, diagnostic, predictive and prescriptive analysis

**Descriptive Analytics** tells you what happened in the past.

Large parts of my report are looking at trends in electrical vehicle, electric vehicle charging and NO2 data to date. These fall into descriptive analysis as it give insights into what has happened in the past.

**Diagnostic Analytics** helps you understand why something happened in the past.

NA

**Predictive Analytics** predicts what is most likely to happen in the future.

I analysised NO2 conc data and used a forecast to look into the future to predict what NO2 conc may look like. This would be predictive analysis as I'm trying to make a prediction of future data.

**Prescriptive Analytics** recommends actions you can take to affect those outcomes.

I analysed electric vehicle charger distribution and affordability of electric vehicles before making a recommendation of actions that could be taken to improve accessibility. This would fall under prescriptive analysis and I have recommended actions to affect outcomes.

### 2021-10-04 documentation_and_analysis/draftbook.Rmd
Edits

@@
   labs(title = "\nEVs vs Mean Annual Income for Local Authorities in England and Wales\n",
-       x = "\nTotal Number of EV Chargers (log)\n",
+       x = "\nNumber of Electric Vehicles (log)\n",
        y = "\nMean Total Annual Income\n") +
